---
title: "การจัดการข้อมูลอนุกรมเวลา ts xts tsbox timetk"
author: "Piyayut Chitchumnong"
date: "2020-10-07"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
---

R ecosystem มีเครื่องมือในการจัดการข้อมูลอนุกรมเวลาที่หลากหลาย มีมากกว่า 14 class แต่ class ที่เป็นที่รู้จักโดยทั่วไปมี  
- `ts` จาก `stats`  
- `zoo`  
- `xts` จาก  
- `tbl`  

แนวคิดในการจัดการข้อมูลอนุกรมเวลาโดยใช้ `data.frame`
tsbox provides a set of tools that are agnostic towards the existing standards. The tools also allow you to handle time series as plain data frames, thus making it easy to deal with time series in a dplyr or data.table workflow.

tsbox is built around a set of converters, which convert time series stored as ts, xts, data.frame, data.table, tibble, zoo, tsibble, tibbletime or timeSeries to each other. Because this works reliably, we can easily write functions that work for all classes. So whether we want to smooth, scale, differentiate, chain, forecast, regularize or seasonally adjust a time series, we can use the same commands to whatever time series class at hand. And, most conveniently, we get a time series plot function that works for all classes and frequencies.

Install package จาก CRAN

```{r}
# install.packages("tsbox")
# install.packages("timetk")
library(tsbox)
library(timetk)
library(tidyverse)
```

# Conversion
```{r}
x.ts <- ts_c(fdeaths, mdeaths)
x.xts <- ts_xts(x.ts)
x.df <- ts_df(x.ts)
x.tbl <- ts_tbl(x.ts)
```

```{r ex, echo=FALSE}
str(x.ts)
```

```{r}
str(x.xts)
```

```{r}
str(x.df)
```

```{r}
str(x.tbl)
```

```{r}
x.ts %>% head()
```

```{r}
ts_scale(x.ts)
```

```{r}
m4_quarterly

ts_ts(m4_quarterly)
```

```{r}
tk_ts(x.df %>% filter(id == "mdeaths"), start = 1974, freq = 12)
```

```{r}

```
